---
title: "Manta Ray Movement Across Indonesia"
subtitle: "Tracking Movement Data"
author: "Norina Haefelin"
institute: "Rosenstiel School of Atmospheric, Earth, and Marine Science"
format: 
  revealjs:
    incremental: true
    code-fold: true
---

## Introduction 
* Manta ray are cute  
* They often make short-range shuffling movements along upwelling events and thermal fronts in Mexico
* Little is known about correlations of their movement in Indonesia, where they are often seen close to shore
* Other satellite tagging of manta rays in Mexico has shown that a majority of the time they are in shallow waters no deeper than 50 m deep (Graham et al., 2012)

## Objectives
* This manta ray tracking data set is looking at these four regions in Indonesia to analyze insight into their spatial movement patterns
* Indonesia is home to significant populations of globally vulnerable reef manta rays in four key regions
* The purpose of the present data set is to analyze Manta ray movements across various areas in Indonesia to analyze difference in movements
* I wanted to analyze manta ray abundance in these four different areas to see correlations/differences in manta ray movement

## Data Sources 
* All tracking data downloaded from MoveBank, an online data platform that allows users to publish animal tracking datasets. All data available to the public

* The data looked at 25 different manta rays across 4 locations in Indonesia (Berau, Nusa Penida, Komodo, Raja Ampat)

* <https://www.datarepository.movebank.org>

## Data Processing 
* The original data was 25 Individual manta rays tracked with 2,960 total observations across 4 different sites. 
* It was a lot of observations.
* To clean up the data, I used longitude and latitude with the function st_as_sf(coords) to make sure the lon and lat stayed with their data.
* Then i used group_by() and summarize(), grouping locality with the individual local identifyer, and summerized number of observations to find observation per manta per site.
* This gave me my final data set to use for data visualization, essentially  25 observations with 4 variables.

## First Analysis
* At first the researcher (me) analyzed correlation between coral reefs and manta ray tracking data 
```{r}
#| echo: false
#| message: false
#| warning: false
# Load packages
library(EVR628tools)
library(tidyverse)
library(mapview)
library(sf)
library(rnaturalearth)
library(terra)
library(ggspatial)
library(tidyterra)

# Load Data
manta_obs_per_site <- read_rds("data/processed/obs_per_manta_site.rds")

reefs <- ne_download(scale = 10L, type = "reefs", category = "physical")

depth <- rast("data/raw/depth_raster.tif")

coast <- ne_countries(country = "Indonesia")

#Map View 
mapview(list(manta_obs_per_site, reefs))
```
# That proved unsuccessful. No significant correlation between reefs and manta ray movement.

## Main Findings 
* So next, the researcher (still me) analyzed manta ray abundance to depth (see next slide) and found that these manta ray do enjoy shallower waters
* These are congurent results found in Mexico

#
```{r}
#| code-fold: true
#| message: false
#| warning: false
#| echo: true
indonesia_depth <- crop(depth, manta_obs_per_site)

indonesia_coast <- st_crop(coast, manta_obs_per_site)

ggplot() +
  geom_spatraster_contour(data = indonesia_depth,
                          aes(colour = after_stat(level)),
                          linewidth = 0.5) +
  geom_sf(data = manta_obs_per_site, color = "red") +
  geom_sf(data = indonesia_coast, fill = "darkgrey", color = "purple") +
  scale_color_viridis_c(option = "mako") +
  annotation_north_arrow(location = "tr", pad_x = unit(.0001, "in"),
        pad_y = unit(.0001, "in")) +
  annotation_scale(location = "br") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(plot.margin = unit(c(1, 1, 1, 1), "in")) +
  coord_sf(clip = "off") +
  theme_classic() +
  labs(title = "Manta Ray Tracking data in Indonesia",
       subtitle = "Analysis of Manta Ray Abundance to Depth",
       x = "Longitutde",
       y = "Latitude",
       color = "Depth (m)",
       caption = "Data: MoveBank Tracking Data")


```

## Final Observations
* Finally, I looked at number of manta rays found at each of the four sites and found significantly higher number of manta rays at Raja Ampat, potentially suggesting a better environment for manta rays

```{r}
#| code-fold: true
#| message: false
#| warning: false
#| echo: true
ggplot(data = manta_obs_per_site,
      mapping = aes(x = locality, y = n_observations)) +
  geom_col(fill = "blue") +
  labs(x = "Location",
       y = "Total Observed",
       subtitle = "Total Number of Manta Observed by Location",
       caption = "Data from DataBank Tracking") +
theme_minimal(base_size = 12)
```


# Thank you for a great semester! :)


## References 

* Graham, R. T., Witt, M. J., Castellanos, D. W., Remolina, F., Maxwell, S., Godley, B. J., & Hawkes, L. A. (2012). Satellite tracking of manta rays highlights challenges to their conservation. *PLoS ONE, 7(5)*, 1-6. <https://doi.org/10.1371/journal.pone.0036834>

* <https://datarepository.movebank.org/entities/datapackage/0425c92c-f157-4f92-88f1-f40de2e0405f>
